---
title: Supplementary Materials
date: last-modified
date-format: long
format: html
toc: true
number-sections: true
author:
  - name: James Steele
    affil-id: 1,2,3
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james@steele-research.com
    affiliations:
      - id: 1
        name: Steele Research Limited, Eastleigh, Hampshire, UK
      - id: 2
        name: MacroFactor, Stronger by Science Technologies LLC, Raleigh, North Carolina, USA
      - id: 3
        name: School of Health and Biomedical Sciences, Royal Melbourne Institute of Technology, Melbourne, VIC, Australia
bibliography: 
  references.bib
csl: apa.csl
execute: 
  message: false
  warning: false
---
```{r}
#| include: false

targets::tar_config_set(store = here::here('_targets'))
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
options(knitr.kable.NA = '')

library(tidyverse)

```

This supplementary material accompanies the preprint: 

> Steele, J. (2026). Little evidence for meaningful additive intervention effect heterogeneity of resistance training upon strength and hypertrophy: Evidence from causal inference and meta-analytic variance comparisons in a large dataset of randomised controlled trials. sportrxiv DOI: [https://doi.org/10.51224/SportRxiv.????](https://doi.org/10.51224/SportRxiv.????). Address for correspondence: james@steele-research.com

# Assumed values of $\rho_{Int:Con}$ for a given variance ratio
As shown by Caldwell et al. [-@caldwellConcernsRegardingStandard2025], the assumed value of $\rho_{Int:Con}$ is dependent upon the ratio between the observed standard deviations in the intervention and control conditions in any variance comparisons performed. This is visualised in @fig-assumed-rho.

```{r}
#| message: false
#| warning: false
#| echo: false

targets::tar_load(rho_assumptions_plot)

```


```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-assumed-rho
#| fig-width: 10
#| fig-height: 5
#| fig-cap: The relationship between $\rho_{Int:Con}$ and the ratio between the observed standard deviations in the intervention and control conditions

rho_assumptions_plot

```

# Mean-variance relationship
This relationship between means and standard deviations on both the raw and log transformed scales for the present dataset is visualised in @fig-mean-var.

```{r}
#| message: false
#| warning: false
#| echo: false

targets::tar_load(mean_var_plot)

```


```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-mean-var
#| fig-width: 10
#| fig-height: 7.5
#| fig-cap: Scatter plots of raw mean and standard deviation of post-intervention scores for (a) strength outcomes and (b) hypertrophy outcomes, and of the log mean and log standard deviation of post-intervention scores for (c) strength outcomes and (d) hypertrophy outcomes.

mean_var_plot

```

# Assumed values of $\rho_{Int:Con}$ in the current dataset for all variance comparisons
According to Caldwell et al. [-@caldwellConcernsRegardingStandard2025], the assumed value of $\rho_{Int:Con}$ was calculated from the ratio between the observed standard deviations in the intervention and control conditions in all variance comparisons performed. This is visualised in @fig-assumed-rho-data.

```{r}
#| message: false
#| warning: false
#| echo: false

targets::tar_load(rho_assumptions_data_plot)

```


```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-assumed-rho-data
#| fig-width: 10
#| fig-height: 7.5
#| fig-cap: Histogram of all assumed values of $\rho_{Int:Con}$ calculated from the observed standard deviations in the intervention and control conditions in all variance comparisons for the current dataset.

rho_assumptions_data_plot

```

# Comparison of between condition (i.e., intervention minus control) and within condition (i.e., post minus pre scores for intervention) average intervention effect estimates
To check whether $\overline{D}=\overline{Y_{Post}}-\overline{Y_{Pre}}$ is  an unbiased estimate of $D=Y_{Int}-Y_{Con}$ I compared across all studies and outcomes the sample based standardised effect size estimate either between intervention and control for a pre-post RCT design (i.e., Using Morris' [-@morrisEstimatingEffectSizes2008] $PPC_2$ with the pooled pre-intervention standard deviation as denominator) and the corresponding standardised effect estimate from the within intervention arm only from pre-intervention to post-intervention (i.e., using Becker's [-@beckerSynthesizingStandardizedMeanchange1988] $d$ and the pre-intervention standard deviation as denominator) with corresponding small sample bias adjustment for both. Both effects were very strongly correlated with no obvious systematic bias between estimates and this visualised can be seen in @fig-between-within.

```{r}
#| message: false
#| warning: false
#| echo: false

targets::tar_load(between_within_ates_plot)

```


```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-between-within
#| fig-width: 7.5
#| fig-height: 5
#| fig-cap: Scatter plot of Morris' [-@morrisEstimatingEffectSizes2008] $PPC_2$ and Becker's [-@beckerSynthesizingStandardizedMeanchange1988] $d$ both with small sample bias adjustment for the current dataset.

between_within_ates_plot

```
